---
output:
  pdf_document: default
  html_document: default
---
___
title: "OnsetFiles: R Package"
output: html_document
___

setwd("~/Desktop/CompStatFiles")

#Function 1: make.datalist
Usability: user provides an input pattern and the file extension. Function will search for all files that fit that description in the working directory and compile them into a list.

```{r}
#### function to read in data ####
make.datalist <- function(pattern, file){
  #makes list of files that meet the pattern
  filelist=list.files(pattern=pattern)
  #drops file extension
  filenames <- gsub('.{4}$','',filelist)
  #puts files into a list object
  if(isTRUE(file=="csv")) {
  d=lapply(filelist, function(x)read.csv(x, header=T)) }
  else if(isTRUE(file=="txt")){   
  d=lapply(filelist, function(x)read.delim(x, header=T)) }
  else {
    print("Warning: data must be txt or csv format")}
  #adds names from filenames to elements of list
  names(d) <- paste(filenames)
  return(d)
}
```

make.datalist works with txt and csv files only. Any other file type will be given a warning.

```{r}
readtest_txt <- make.datalist("*Enc1.txt", "txt")
readtest_csv <- make.datalist("*Enc1.csv","csv")
readtest_xlsx <- make.datalist("*Enc1.xlsx", "xlsx")
```


#Function 2: make.duration
Usability: user provides the dataset (a list) and the names of two columns. Column1 will be subtracted from Column2. This number will be transformed from scanner time to seconds, then put into a list. When calculating the durations, negative values will be changed to NAs. If this happens, the user will see a warning that negative values were found.

```{r}
#### function for making duration ####
#only used when you want to subtract one column from another!
make.duration <- function(data, col1, col2){
  duration_list <- vector(mode="list", length=length(data))
  for(i in 1:length(data)){
    #make vector of durations
    duration <- (data[[i]][col2] - data[[i]][col1])/1000
    #check for negative values, return "NA" in place of negatives
    omitneg <- function(x){ifelse(x< 0, "NA",x)}
    dur_omit <- sapply(duration, omitneg)
    #check for NAs, and return warning if any NAs found
    if(isTRUE(all.equal(duration,dur_omit))){
    } else {
      print(paste(i, "Warning: Negative values calculated and set to NA"))
    }
    duration_list[[i]] <- as.numeric(dur_omit)
    #name each column based on iteration of data list
    names(duration_list) <- names(data)
  }
  return(duration_list)
}
```


A warning will be printed for each subject that has NAs in their durations.
```{r}
durationtest <- vector(mode="list", length = 2)
durationtest <- make.duration(readtest_csv, "Vividness.OnsetTime", "Vividness.RTTime")
```


#Function 3: make.onset
Usability: user provides the data (a list), the name of the scanner trigger column, the condition of interest for the onset file, the duration (a list or single number), and the weight (a single number). The function will output a 3 column dataframe of Onset, Duration, and Weight in the format of a list.

```{r}
#### function for onset files ####
make.onset <- function(data, trigger, condition, duration, weight) {
  onset_file <- vector(mode="list", length(data))
  onset_file_omit <- vector(mode="list", length(data))
  for(i in 1:length(data)){
  #make onset column: put from scanner time into real time
  cond_onset <- (data[[i]][trigger] - data[[i]][condition])/1000
  #make duration column
  if(isTRUE(is.numeric(duration))){
  #if argument is a number, repeat for length of onset file
  dur <- rep(duration, length(cond_onset))
  }
  else if(isTRUE(is.list(duration))){
  #if argument is a list, copy variable to onset file
  dur <- duration[[i]]
  }
  else{
  print(paste(i, "Warning: Duration must be specified as a single number or list"))
  }
  #make weight column
  wgt <- rep(weight, length(cond_onset))
  #put together in a dataframe
  onset_file[[i]] <- data.frame(cond_onset, dur, wgt)
  #omit rows with value NA
  onset_file_omit[[i]] <- na.omit(onset_file[[i]])
  if(isTRUE(all.equal(onset_file_omit[[i]],onset_file[[i]]))){
  } else {
  print(paste(i, "Warning: NAs were removed from onset file"))
  }
  }
  #name each column based on iteration of data list
  names(onset_file_omit) <- names(data)
  #name each column in each list element
  colnames <- c("Onset", "Duration", "Weight")
  onset_file_done <- lapply(onset_file_omit, setNames, colnames)
  return(onset_file_done)
}
```

make.onset will determine whether a list or single value was entered for duration. If another input is given, the user will see a warning. The function will omit from the final dataframe any rows with a value of NA, and the user will be warned.

With a constant duration:
```{r}
onsettest_singledur <- vector(mode="list", length=length(readtest_csv))
onsettest_singledur <- make.onset(readtest_csv, "Trigger.RTTime", "Fixation.OnsetTime", 2, 1)
```

With a variable duration:
```{r}
onsettest_vardur <- vector(mode="list", length=length(readtest_csv))
onsettest_vardur <- make.onset(readtest_csv, "Trigger.RTTime", "Fixation.OnsetTime", durationtest, 1)
```

The user can then save these onset files as txt files using their preferred method.
